\section{mbeddr Debugger}

mbeddr comes with an extensible debugger, that allows user to debug code written
with mbeddr and which can be extended to support debugging of new language
extensions.

\subsection{Architecture}
how does the debugger work

Execution Aspect

Specification Aspect
DSL


\subsection{Building Debugging Behavior}

We describe in this section the implementation of debugging behavior for the
lanuages described in \sect{languageImplementation}. The mbeddr debugger
comes with a \ac{DSL} for specifying the debugging behavior. 
We will use the language in this section for
building debug support for both languages from \sect{languageImplementation}.

\parhead{Breakpoints} We want to be able to set breakpoints on instances of
\ic{ForEachStatement}, \ic{AssertStatement}, \ic{ItExpression}and
\ic{ExecuteTestExpression}.
For doing so, we need to implement the marker interface \ic{IBreakable}. 
Since the first two concepts
are derived from \ic{Statement} and this concept already implements this
interface, we can already set breakpoints on instances of them. The latter two
concepts will always live inside a statement derived
from \ic{Statement}. Hence, no further work is necessary.

\parhead{Call Stack} For lifting the call stack we need to specify, which
concepts have callable semantics or are generated to C functions. First applies
to \ic{TestCase}, which has callable semantics, but is also trasnalted to a C
function. In contrast, \ic{ExecuteTestExpression} does not have callable
semantics, but is translated to a C function. Other concepts are not
related to the call stack.

\ic{TestCase} , \ic{ExecuteTestExpression} \ic{IStackFrameContributor}






\parhead{Stepping}

\parhead{Watches}
